---
title: "Bin_the_genome_keeping strand"
author: "Amalia"
date: "2023-10-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Bin the CpG files into 1kb/10kb sized windows
We want to bin our genome into tiles, and get the ratio of methylation in purines and pyrimidines in each of these subgroups.
In order To do that we have to calculate the average methylation of each CGN, and we will also count the number of mapped CGNs, which is what this script will do.
We will keep the strand 

load the libraries
```{r, message=FALSE}
library(GenomicRanges)
library(tidyverse)
library (purrr)
library(plyranges)
library(rlist)
library(BSgenome.Mmusculus.UCSC.mm9)
```

As a first step, we have to divide the CpGs into the group. We do so by "annotating" each CpG with the number of (genomic) tile it falls into.
```{r}
#we get the tiles by cutting the mm9 genome 
tiles<-tileGenome(seqinfo(Mmusculus),tilewidth = 1000, cut.last.tile.in.chrom = TRUE)

tile.df<-as.data.frame(tiles) #used later 

#add strand


#
tileplus_df<-tile.df
tile_plus_df<-tile.df %>% mutate(strand = "+")
tile_minus_df<-tile.df %>% mutate(strand = "-")
tile_plus<-tile_plus_df %>% makeGRangesFromDataFrame()
tile_minus<-tile_minus_df %>% makeGRangesFromDataFrame()

seqinfo(tile_minus)<-Seqinfo(genome = "mm9")
#test
#a_3<-GenomicRanges::trim(tile_minus) #because we might get out of bounds
```
We will first bin each strand separately (basically we will use either + or - for overlapping) and we will name the bin accordingly (so where we had bin 1000, now we will have bin 1000_plus and bin 1000_minus)
get the average meth in each bin and total number 
```{r}
#create a vector holding the names of all the sample names

sample_names<-c("GSM4594635","GSM1382253","GSM1382254","GSM2533056","GSM4837324","GSM1382255","GSM1382256","GSM1545748","GSM748786","GSM748787")
for(i in 2:10){
  a<-readRDS(file = paste("../data/raw_data/CpG/", sample_names[i],"_mCpG.rds" ,sep=""))
  a<-a%>% as.data.frame() %>% `colnames<-`(c("seqnames", "start","end","width","strand","T","M", "CGN_seq", "score")) %>% makeGRangesFromDataFrame(keep.extra.columns = T) 
  a<- a[a$T>4 & a$T<101] #filter for at least 5 reads, to have enough points (10 is too strict)
  #add the bins in the sample = information about which tile it belongs to 
  ah_plus<-findOverlaps(a,tile_plus) #for plus strand
  ah_minus<-findOverlaps(a,tile_minus) #for minus strand
  ##we will add column about the "position" (in which 1kb bin it falls into)
  mcols(a)$bin<-NA
  #mcols(a)$bin[queryHits(ah)]<-subjectHits(ah)
  mcols(a)$bin[queryHits(ah_plus)]<-subjectHits(ah_plus)
  mcols(a)$bin[queryHits(ah_minus)]<-subjectHits(ah_minus)
  ## by doing tiles[ bin number] you can find the bin 
  a<-a %>% mutate(bin = ifelse(strand=="+", paste0(bin, "+"), paste0(bin, "-"))) #add info about which strand this bin is from
  d<-as.data.frame(a)
  
  #get average meth score per bin and number of cpgs
  avg<- d %>% group_by(bin) %>% summarise(avg_meth=mean(score))
  tot<- d %>% group_by(bin) %>% summarise(CG_num=n())
  totl<-merge(avg, tot, by="bin")
  #for arranging
  totl<-totl %>% mutate(bin_only_number= str_sub(bin, 1,nchar(bin)-1)) %>% mutate_at('bin_only_number', as.integer)
  totl<-totl %>% arrange(bin_only_number)
  
  #get average meth score per CGN per bin
  e<-d %>% group_by(bin,CGN_seq) %>% summarise(mean_meth=mean(score))
  e<-e %>% spread(CGN_seq,mean_meth)
  
  d<- d %>% mutate(pur_pyr = case_when(CGN_seq == "CGA" | CGN_seq == "CGG" ~"Purine", CGN_seq == "CGT" | CGN_seq == "CGC" ~ "Pyrimidine"))
  e_p<-d %>% group_by(bin,pur_pyr) %>% summarise(mean_meth=mean(score))
  e_p<-e_p %>% spread(pur_pyr,mean_meth)
  e<-merge(e, e_p, by="bin")
  #arraneg correctly
  e<-e %>% mutate(bin_only_number= str_sub(bin, 1,nchar(bin)-1)) %>% mutate_at('bin_only_number', as.integer)
  e<-e %>% arrange(bin_only_number)
  
  colnames(e)<-paste(colnames(e), "meth",sep="_" )
  
  #get number of each CGN per bin
  e_2<-d %>% group_by(bin, CGN_seq) %>% summarise(count=n(), .groups = 'drop')
  e_2<-e_2 %>% spread(CGN_seq,count)
  e_p2<-d %>% group_by(bin,pur_pyr) %>% summarise(count=n(), .groups = 'drop')
  e_p2<-e_p2 %>% spread(pur_pyr,count)
  e_2<-merge(e_2, e_p2, by="bin")
  colnames(e_2)<-paste(colnames(e_2), "num", sep = "_")
  e_2<-e_2 %>% mutate(bin_only_number= str_sub(bin_num, 1,nchar(bin_num)-1)) %>% mutate_at('bin_only_number', as.integer) #get everything except the last , so only number
  e_2<-e_2 %>% mutate(bin_strand= str_sub(bin_num, start= -1)) #get only last, so strand
  e_2<-e_2 %>% arrange(bin_only_number)#reorder so that we have normal order of bins
 
  df<-cbind(e,e_2)
  df<-cbind(totl,df)
  df_2<-df %>% select(unique(colnames(.)))
  df_2<-df_2 %>% select(-c("bin_num", "bin_meth")) 
  pf<-tile.df %>% mutate(rn= row_number()) %>% filter(rn %in% e_2$bin_only_number) %>% rename(bin_only_number=rn) #twice
  df_2<-left_join(df_2,pf, by="bin_only_number")
  #mutate teh strand and change it again to factor
  df_2<-df_2 %>% mutate(new_strand = as.factor(bin_strand)) %>% select(-c("strand", "bin_strand")) %>% rename(strand=new_strand)
  df_2<-df_2 %>% select(- "bin_only_number_meth")
  saveRDS(df_2, file=paste0("../data/processed/bins_with_strand_information/CpGs/1kb/", sample_names[i], ".RDS"))
}
#df_2 %>% mutate(new_strand = as.factor(bin_strand)) %>% select(-c("strand", "bin_strand")) %>% rename(bin=bin_meth) %>% rename(strand=new_strand)
```
same for 10 kb
```{r}
#we get the tiles by cutting the mm9 genome 
tiles<-tileGenome(seqinfo(Mmusculus),tilewidth = 10000, cut.last.tile.in.chrom = TRUE)
#DNMT3A<-readRDS("../methylation_project/data/TKO_DNMT3A2_wt_i12_WGBS_CGN.gr.rds")
tile.df<-as.data.frame(tiles) #used later 

```
Average meth and n of CpGs 
```{r}
#create a vector holding the names of all the sample names

sample_names<-c("GSM4594635","GSM1382253","GSM1382254","GSM2533056","GSM4837324","GSM1382255","GSM1382256","GSM1545748","GSM748786","GSM748787")
for(i in 1:10){
  a<-readRDS(file = paste("../data/raw_data/CpG/", sample_names[i],"_mCpG.rds" ,sep=""))
  a<-a%>% as.data.frame() %>% `colnames<-`(c("seqnames", "start","end","width","strand","T","M", "CGN_seq", "score")) %>% makeGRangesFromDataFrame(keep.extra.columns = T) 
  a<- a[a$T>4 & a$T<101] #filter for at least 5 reads, to have enough points (10 is too strict)
  #add the bins in the sample = information about which tile it belongs to 
  ah<-findOverlaps(a,tiles)
  ##we will add column about the "position" (in which 1kb bin it falls into)
  mcols(a)$bin<-NA
  mcols(a)$bin[queryHits(ah)]<-subjectHits(ah)
  ## by doing tiles[ bin number] you can find the bin 
  d<-as.data.frame(a)
  
  #get average meth score per bin and number of cpgs
  avg<- d %>% group_by(bin) %>% summarise(avg_meth=mean(score))
  tot<- d %>% group_by(bin) %>% summarise(CG_num=n())
  totl<-merge(avg, tot, by="bin")
  
  #get average meth score per CGN per bin
  e<-d %>% group_by(bin,CGN_seq) %>% summarise(mean_meth=mean(score))
  e<-e %>% spread(CGN_seq,mean_meth)
  
  d<- d %>% mutate(pur_pyr = case_when(CGN_seq == "CGA" | CGN_seq == "CGG" ~"Purine", CGN_seq == "CGT" | CGN_seq == "CGC" ~ "Pyrimidine"))
  e_p<-d %>% group_by(bin,pur_pyr) %>% summarise(mean_meth=mean(score))
  e_p<-e_p %>% spread(pur_pyr,mean_meth)
  e<-merge(e, e_p, by="bin")
  
  colnames(e)<-paste(colnames(e), "meth",sep="_" )
  
  #get number of each CGN per bin
  e_2<-d %>% group_by(bin, CGN_seq) %>% summarise(count=n(), .groups = 'drop')
  e_2<-e_2 %>% spread(CGN_seq,count)
  e_p2<-d %>% group_by(bin,pur_pyr) %>% summarise(count=n(), .groups = 'drop')
  e_p2<-e_p2 %>% spread(pur_pyr,count)
  e_2<-merge(e_2, e_p2, by="bin")
  colnames(e_2)<-paste(colnames(e_2), "num", sep = "_")
  
  df<-cbind(e,e_2)
  df<-cbind(totl,df)
  df<-df %>% select(-c("bin_num","bin_meth")) 
  pf<-tile.df %>% mutate(rn= row_number()) %>% filter(rn %in% e_2$bin_num)
  df<-cbind(df, pf) %>% select(-"rn")
  saveRDS(df, file=paste0("../data/processed/bins/CpGs/10kb/", sample_names[i], ".RDS"))
}

```
#Now lets bin the genomes for the CHH data
#Bin the CHH files into 1kb/10kb sized windows
We want to bin our genome into tiles, and get the ratio of methylation in purines and pyrimidines in each of these subgroups.
In order To do that we have to calculate the average methylation of each CHH, and we will also count the number of mapped CHHs, which is what this script will do.

As a first step, we have to divide the CpGs into the group. We do so by "annotating" each CpG with the number of (genomic) tile it falls into.
```{r}
#we get the tiles by cutting the mm9 genome 
tiles<-tileGenome(seqinfo(Mmusculus),tilewidth = 1000, cut.last.tile.in.chrom = TRUE)
#DNMT3A<-readRDS("../methylation_project/data/TKO_DNMT3A2_wt_i12_WGBS_CGN.gr.rds")
tile.df<-as.data.frame(tiles) #used later 

```
test
filter based on number of reads
remove CGs
```{r}
i=1
j=21
a<-readRDS(file = paste0("../data/raw_data/CHH/", sample_names[i],"_", Total_chr[j],".rds"))
a<- a[a$T>4 & a$T<101] #filter for at least 5 reads, to have enough points (10 is too strict)
  #add the bins in the sample = information about which tile it belongs to
a<-filter(a, !CGN_seq %in% c("CGA", "CGC","CGG","CGT")) #filter out CGs
#unique(a$CGN_seq) sanity check that no CGs exist anymore
ah<-findOverlaps(a,tiles)
  ##we will add column about the "position" (in which 1kb bin it falls into)
  mcols(a)$bin<-NA
  mcols(a)$bin[queryHits(ah)]<-subjectHits(ah)
  ## by doing tiles[ bin number] you can find the bin 
  d<-as.data.frame(a)
  
  #get average meth score per bin and number of cpgs
  avg<- d %>% group_by(bin) %>% summarise(avg_meth=mean(score))
  tot<- d %>% group_by(bin) %>% summarise(CG_num=n())
  totl<-merge(avg, tot, by="bin")
  
  #get average meth score per CHH per bin
  e<-d %>% group_by(bin,CGN_seq) %>% summarise(mean_meth=mean(score))
  e<-e %>% spread(CGN_seq,mean_meth)
  
  d<- d %>% mutate(pur_pyr = case_when(CGN_seq == "CGA" | CGN_seq == "CGG" ~"Purine", CGN_seq == "CGT" | CGN_seq == "CGC" ~ "Pyrimidine"))
  e_p<-d %>% group_by(bin,pur_pyr) %>% summarise(mean_meth=mean(score))
  e_p<-e_p %>% spread(pur_pyr,mean_meth)
  e<-merge(e, e_p, by="bin")
  
  colnames(e)<-paste(colnames(e), "meth",sep="_" )
  
  #get number of each CGN per bin
  e_2<-d %>% group_by(bin, CGN_seq) %>% summarise(count=n(), .groups = 'drop')
  e_2<-e_2 %>% spread(CGN_seq,count)
  e_p2<-d %>% group_by(bin,pur_pyr) %>% summarise(count=n(), .groups = 'drop')
  e_p2<-e_p2 %>% spread(pur_pyr,count)
  e_2<-merge(e_2, e_p2, by="bin")
  colnames(e_2)<-paste(colnames(e_2), "num", sep = "_")
  
  df<-cbind(e,e_2)
  df<-cbind(totl,df)
  df<-df %>% select(-c("bin_num","bin_meth")) 
  pf<-tile.df %>% mutate(rn= row_number()) %>% filter(rn %in% e_2$bin_num)
  df<-cbind(df, pf) %>% select(-"rn")
```
test
```{r}
#create a vector holding the names of all the sample names

sample_names<-c("GSM4594635","GSM1382253","GSM1382254","GSM2533056","GSM4837324","GSM1382255","GSM1382256","GSM1545748","GSM748786","GSM748787")
Total_chr  <-  c(paste("chr", 1:19, sep=''), 'chrX','chrY')
for(i in 1:10){
  for(j in 1:21){
  a<-readRDS(file = paste0("../data/raw_data/CHH/", sample_names[i],"_", Total_chr[j],".rds"))
  a<- a[a$T>4 & a$T<101] #filter for at least 5 reads, to have enough points (10 is too strict)
  #add the bins in the sample = information about which tile it belongs to 
  a<-filter(a, !CGN_seq %in% c("CGA", "CGC","CGG","CGT")) #filter out CGs
 #unique(a$CGN_seq) sanity check that no CGs exist anymore
  ah<-findOverlaps(a,tiles)
  ##we will add column about the "position" (in which 1kb bin it falls into)
  mcols(a)$bin<-NA
  mcols(a)$bin[queryHits(ah)]<-subjectHits(ah)
  ## by doing tiles[ bin number] you can find the bin 
  d<-as.data.frame(a)
  
  #get average meth score per bin and number of cpgs
  avg<- d %>% group_by(bin) %>% summarise(avg_meth=mean(score))
  tot<- d %>% group_by(bin) %>% summarise(CG_num=n())
  totl<-merge(avg, tot, by="bin")
  
  #get average meth score per CGN per bin
  e<-d %>% group_by(bin,CGN_seq) %>% summarise(mean_meth=mean(score))
  e<-e %>% spread(CGN_seq,mean_meth)
  
  d<- d %>% mutate(pur_pyr = case_when(CGN_seq == "CGA" | CGN_seq == "CGG" ~"Purine", CGN_seq == "CGT" | CGN_seq == "CGC" ~ "Pyrimidine"))
  e_p<-d %>% group_by(bin,pur_pyr) %>% summarise(mean_meth=mean(score))
  e_p<-e_p %>% spread(pur_pyr,mean_meth)
  e<-merge(e, e_p, by="bin")
  
  colnames(e)<-paste(colnames(e), "meth",sep="_" )
  
  #get number of each CGN per bin
  e_2<-d %>% group_by(bin, CGN_seq) %>% summarise(count=n(), .groups = 'drop')
  e_2<-e_2 %>% spread(CGN_seq,count)
  e_p2<-d %>% group_by(bin,pur_pyr) %>% summarise(count=n(), .groups = 'drop')
  e_p2<-e_p2 %>% spread(pur_pyr,count)
  e_2<-merge(e_2, e_p2, by="bin")
  colnames(e_2)<-paste(colnames(e_2), "num", sep = "_")
  
  df<-cbind(e,e_2)
  df<-cbind(totl,df)
  df<-df %>% select(-c("bin_num","bin_meth")) 
  pf<-tile.df %>% mutate(rn= row_number()) %>% filter(rn %in% e_2$bin_num)
  df<-cbind(df, pf) %>% select(-"rn")
  saveRDS(df, file=paste0("../data/processed/bins/CpGs/1kb/", sample_names[i], ".RDS"))
  }
}

```

get the average meth in each bin and total number 
```{r}
#create a vector holding the names of all the sample names

sample_names<-c("GSM4594635","GSM1382253","GSM1382254","GSM2533056","GSM4837324","GSM1382255","GSM1382256","GSM1545748","GSM748786","GSM748787")
for(i in 1:10){
  for(j in 1:21){
  a<-readRDS(file = paste0("../data/raw_data/CHH/", sample_names[i],"_", Total_chr[j],".rds"))
  a<-a%>% as.data.frame() %>% `colnames<-`(c("seqnames", "start","end","width","strand","T","M", "CGN_seq", "score")) %>% makeGRangesFromDataFrame(keep.extra.columns = T) 
  a<- a[a$T>4 & a$T<101] #filter for at least 5 reads, to have enough points (10 is too strict)
  #add the bins in the sample = information about which tile it belongs to 
  ah<-findOverlaps(a,tiles)
  ##we will add column about the "position" (in which 1kb bin it falls into)
  mcols(a)$bin<-NA
  mcols(a)$bin[queryHits(ah)]<-subjectHits(ah)
  ## by doing tiles[ bin number] you can find the bin 
  d<-as.data.frame(a)
  
  #get average meth score per bin and number of cpgs
  avg<- d %>% group_by(bin) %>% summarise(avg_meth=mean(score))
  tot<- d %>% group_by(bin) %>% summarise(CG_num=n())
  totl<-merge(avg, tot, by="bin")
  
  #get average meth score per CGN per bin
  e<-d %>% group_by(bin,CGN_seq) %>% summarise(mean_meth=mean(score))
  e<-e %>% spread(CGN_seq,mean_meth)
  
  d<- d %>% mutate(pur_pyr = case_when(CGN_seq == "CGA" | CGN_seq == "CGG" ~"Purine", CGN_seq == "CGT" | CGN_seq == "CGC" ~ "Pyrimidine"))
  e_p<-d %>% group_by(bin,pur_pyr) %>% summarise(mean_meth=mean(score))
  e_p<-e_p %>% spread(pur_pyr,mean_meth)
  e<-merge(e, e_p, by="bin")
  
  colnames(e)<-paste(colnames(e), "meth",sep="_" )
  
  #get number of each CGN per bin
  e_2<-d %>% group_by(bin, CGN_seq) %>% summarise(count=n(), .groups = 'drop')
  e_2<-e_2 %>% spread(CGN_seq,count)
  e_p2<-d %>% group_by(bin,pur_pyr) %>% summarise(count=n(), .groups = 'drop')
  e_p2<-e_p2 %>% spread(pur_pyr,count)
  e_2<-merge(e_2, e_p2, by="bin")
  colnames(e_2)<-paste(colnames(e_2), "num", sep = "_")
  
  df<-cbind(e,e_2)
  df<-cbind(totl,df)
  df<-df %>% select(-c("bin_num","bin_meth")) 
  pf<-tile.df %>% mutate(rn= row_number()) %>% filter(rn %in% e_2$bin_num)
  df<-cbind(df, pf) %>% select(-"rn")
  saveRDS(df, file=paste0("../data/processed/bins/CpGs/1kb/", sample_names[i], ".RDS"))
  }
}

```


```{r}
readRDS("../data/processed/bins_with_strand_information/CpGs/1kb/GSM4594635.RDS")
```

```{r}

```

