---
title: "Examine binned samples"
author: "Amalia"
date: "2023-07-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(GenomicRanges)
library(tidyverse)
library (purrr)
library(plyranges)
library(rlist)
```


From the binned genome, we want to find regions of interest. 

We will rank the tiles according to one characteristic. These characheristics could be:
average methylation of the bin, proportion of methylation in the bin explained by purines/ pyrimidines (we will chose according to the preference found in the whole genome), and "normalized" ratio of methylation in purines over pyrimidines 


```{r}
sample_names<-c("GSM4594635","GSM1382253","GSM1382254","GSM2533056","GSM4837324","GSM1382255","GSM1382256","GSM1545748","GSM748786","GSM748787")
addback<-c("DNMT3A2", "DNMT3A2", "DNMT3A2", "DNMT3A1", "DNMT3B1","DNMT3B1","DNMT3B1","DNMT3B1","WT","WT") #I have to use these as names as well
```
For average methylation
We want to calculate the average methylation on each bin and get the bins with the highest rank (so regions with higher methylation)
```{r}
for(i in 1:10){
  c<-readRDS(file=paste0("../data/processed/bins/CpGs/1kb/", sample_names[i], ".RDS"))
  #We are going to filter out the bins that contain less than 3 mapped CG 
  c<-c %>% filter(CG_num>4)
  c$rank<-rank(-c$avg_meth)#Rank references the position of the value in the sorted vector and is in the same order as the original sequence
  d<-c %>% arrange(rank) #arraange according to rank (we used rank first and not justy arrange on our values immediatelly cause we have a lot of rows with the exact same value)
# Now we want to exract x number of bins with highest rank. Lets say 5%
#to get that 5% we are going to count the number of rows and use 0.05*that and keep the first rows
  t<-round(nrow(c)*0.05) # find n for top 5 %
  e<-head(d,t)
  assign(sample_names[i], e) #gives variable name
}

#Now we have all the dataframes, I will only keep the top bins, to see whether there is overlap between samples
mu_list<-list() #empty list
# we are going to paste there all the top bins to use later with ggven 

for(i in 1:10){
a<-get(sample_names[i]) %>% select(bin)#"get" calls the data by the variable name
colnames(a)<-sample_names[i] 
a <-as.vector(a)
mu_list[sample_names[i]]<-a

}


### for plotting
library(ggvenn)
ggvenn(mu_list[c(1,2,4)], stroke_size = 0.5, set_name_size = 2.5)
ggvenn(mu_list[1:4], stroke_size = 0.5, set_name_size = 2.5)
ggvenn(mu_list[5:8], stroke_size = 0.5, set_name_size = 2.5)
ggvenn(mu_list[9:10], stroke_size = 0.5, set_name_size = 2.5)
ggvenn(mu_list[c(5,7,8)], stroke_size = 0.5, set_name_size = 2.5)

ggvenn(mu_list[c(1,2,5)], stroke_size = 0.5, set_name_size = 2.5)

laal<-ggvenn(mu_list)
attributes(laal)
inter<-euler(mu_list)
plot(inter)
attributes(inter)
inter$original.values
```
For purine/pyrimidine methylation
We want to calculate the purine methylation / pyrimidine methylation ratio in each bin
We will filter out bins that have only mapped CGs followed by purines or only by purines.
We will create an artificial column where we add +0.01 on purine and pyrimidine methylation 
Because we would otherwise (because of the many 0 values ) we would get a lot of ratios that are equal to 0 or infinity.

We will have a column with the log2 transformation of that ratio because that way, a ratio can be comparable when its reversed {eg the ratio of 0.8/0.2 vs 0.2/0.8 would have an equal absolute value of 2 whereas the values without log2 are 4 and 0.25}
```{r}
# I will not do it for the WT cells now
for(i in 1:8){
  c<-readRDS(file=paste0("../data/processed/bins/CpGs/1kb/", sample_names[i], ".RDS"))
  #We are going to filter out the bins that have NA at pur number or pyrimidine number
  c<-c %>% filter(!is.na(Purine_num) & !is.na(Pyrimidine_num))
  c<- c %>% filter(Purine_num>1 & Pyrimidine_num>1) # if we are being strict
  #now create columns with modified methylation
  c<-c %>% mutate(Purine_meth_mod = Purine_meth + 0.01)
  c<-c %>% mutate(Pyrimidine_meth_mod = Pyrimidine_meth + 0.01)
  #now create col with ratio of those two
  c <-c %>% mutate(ratio_pur_pyr = Purine_meth_mod / Pyrimidine_meth_mod)
  #now create col of log2 of that ratio 
  options(scipen = 999)# to avoid scientific notation
  c<-c %>% mutate(ratio_purpyr_log = log2(ratio_pur_pyr)) 
  c$rank<-rank(-c$ratio_purpyr_log)#Rank references the position of the value in the sorted vector and is in the same order as the original sequence 
  #most negative will be rank1 
  #negatives have a pyrimidine preference 
  # arrange differently for A and B, we want to get the most negatives for 3A and the most positives for 3B
  t<-round(nrow(c)*0.05)# find n for top 5 %
  print(t)
  if(i<=4){
    d<-c %>% arrange(-rank)
    e<-head(d,t)
    assign(sample_names[i], e) 
    #get the ones with biggest preference for pyrim (for 3A)
  }else{
    d<-c %>% arrange(rank)
    e<-head(d,t)
    assign(sample_names[i], e)
    #get the ones with biggest preference for pyrim (for 3B)
  }
}
GSM1382253  
#Now we have all the dataframes, I will only keep the top bins, to see whether there is overlap between samples
mu_list_2<-list() #empty list
# we are going to paste there all the top bins to use later with ggven 

for(i in 1:8){
a<-get(sample_names[i]) %>% select(bin)#"get" calls the data by the variable name
colnames(a)<-sample_names[i] 
a <-as.vector(a)
mu_list_2[sample_names[i]]<-a

}
  
### for plotting
library(ggvenn)
ggvenn(mu_list_2[c(1,2,4)], stroke_size = 0.5, set_name_size = 2.5)
ggvenn(mu_list_2[1:4], stroke_size = 0.5, set_name_size = 2.5)
ggvenn(mu_list_2[5:8], stroke_size = 0.5, set_name_size = 2.5)
ggvenn(mu_list_2[c(5,7,8)], stroke_size = 0.5, set_name_size = 2.5)



ggvenn(mu_list_2[c(1,2,8)], stroke_size = 0.5, set_name_size = 2.5)

laal<-ggvenn(mu_list)
attributes(laal)
inter<-euler(mu_list)
plot(inter)
attributes(inter)
inter$original.values
#GSM1382253

#library(VennDiagram)

library(eulerr)
inter<-euler(mu_list_2[c(1,2,4)])
plot(inter, quantities = TRUE)
ggvenn(mu_list[c(1,2,4)], stroke_size = 0.5, set_name_size = 2.5)
error_plot(inter)
```
get the ones with preference for pyrimidine
```{r}
for(i in 1:8){
  c<-readRDS(file=paste0("../data/processed/bins/CpGs/1kb/", sample_names[i], ".RDS"))
  #We are going to filter out the bins that have NA at pur number or pyrimidine number
  c<-c %>% filter(!is.na(Purine_num) & !is.na(Pyrimidine_num))
  c<- c %>% filter(Purine_num>1 & Pyrimidine_num>1) # if we are being strict
  #now create columns with modified methylation
  c<-c %>% mutate(Purine_meth_mod = Purine_meth + 0.01)
  c<-c %>% mutate(Pyrimidine_meth_mod = Pyrimidine_meth + 0.01)
  #now create col with ratio of those two
  c <-c %>% mutate(ratio_pur_pyr = Purine_meth_mod / Pyrimidine_meth_mod)
  #now create col of log2 of that ratio 
  options(scipen = 999)# to avoid scientific notation
  c<-c %>% mutate(ratio_purpyr_log = log2(ratio_pur_pyr)) 
  c$rank<-rank(-c$ratio_purpyr_log)#Rank references the position of the value in the sorted vector and is in the same order as the original sequence 
  #most negative will be rank1 
  #negatives have a pyrimidine preference 
  # arrange differently for A and B, we want to get the most negatives for 3A and the most positives for 3B
  t<-round(nrow(c)*0.05)# find n for top 5 %
  print(t)
  d<-c %>% arrange(-rank)
  e<-head(d,t)
  assign(sample_names[i], e)
  
}
GSM1382253  
#Now we have all the dataframes, I will only keep the top bins, to see whether there is overlap between samples
mu_list_3<-list() #empty list
# we are going to paste there all the top bins to use later with ggven 

for(i in 1:8){
a<-get(sample_names[i]) %>% select(bin)#"get" calls the data by the variable name
colnames(a)<-sample_names[i] 
a <-as.vector(a)
mu_list_3[sample_names[i]]<-a

}
ggvenn(mu_list_2[c(1,2,4)], stroke_size = 0.5, set_name_size = 2.5)
ggvenn(mu_list_2[1:4], stroke_size = 0.5, set_name_size = 2.5)
ggvenn(mu_list_2[5:8], stroke_size = 0.5, set_name_size = 2.5)
ggvenn(mu_list_2[c(5,7,8)], stroke_size = 0.5, set_name_size = 2.5)



ggvenn(mu_list_2[c(2,4,8)], stroke_size = 0.5, set_name_size = 2.5)
```
for purines
```{r}
for(i in 1:8){
  c<-readRDS(file=paste0("../data/processed/bins/CpGs/1kb/", sample_names[i], ".RDS"))
  #We are going to filter out the bins that have NA at pur number or pyrimidine number
  c<-c %>% filter(!is.na(Purine_num) & !is.na(Pyrimidine_num))
  c<- c %>% filter(Purine_num>1 & Pyrimidine_num>1) # if we are being strict
  #now create columns with modified methylation
  c<-c %>% mutate(Purine_meth_mod = Purine_meth + 0.01)
  c<-c %>% mutate(Pyrimidine_meth_mod = Pyrimidine_meth + 0.01)
  #now create col with ratio of those two
  c <-c %>% mutate(ratio_pur_pyr = Purine_meth_mod / Pyrimidine_meth_mod)
  #now create col of log2 of that ratio 
  options(scipen = 999)# to avoid scientific notation
  c<-c %>% mutate(ratio_purpyr_log = log2(ratio_pur_pyr)) 
  c$rank<-rank(-c$ratio_purpyr_log)#Rank references the position of the value in the sorted vector and is in the same order as the original sequence 
  #most negative will be rank1 
  #negatives have a pyrimidine preference 
  # arrange differently for A and B, we want to get the most negatives for 3A and the most positives for 3B
  t<-round(nrow(c)*0.05)# find n for top 5 %
  print(t)
  d<-c %>% arrange(rank)
  e<-head(d,t)
  assign(sample_names[i], e)
  
}
GSM1382253  
#Now we have all the dataframes, I will only keep the top bins, to see whether there is overlap between samples
mu_list_4<-list() #empty list
# we are going to paste there all the top bins to use later with ggven 

for(i in 1:8){
a<-get(sample_names[i]) %>% select(bin)#"get" calls the data by the variable name
colnames(a)<-sample_names[i] 
a <-as.vector(a)
mu_list_4[sample_names[i]]<-a

}
ggvenn(mu_list_2[c(1,2,4)], stroke_size = 0.5, set_name_size = 2.5)
ggvenn(mu_list_2[1:4], stroke_size = 0.5, set_name_size = 2.5)
ggvenn(mu_list_2[5:8], stroke_size = 0.5, set_name_size = 2.5)
ggvenn(mu_list_2[c(5,7,8)], stroke_size = 0.5, set_name_size = 2.5)



ggvenn(mu_list_2[c(1,4,8)], stroke_size = 0.5, set_name_size = 2.5)
```



Now, we are going to calculate the ratio of methylation of CpGs flanked by purines (CGR) over the methylation of CpGs flanked by pyrimidines (CGY). 
Firstly we will calculate the same ratio but for mapped CGR/ CGY in each bin, to see if we have a normal distribution. 


TEST to create list that holds all the bins that are overlapping between all combinations
```{r}
e<-c(1,2,3,4,5,6,7)
f<-c(1,4,8)
g<-c(1,2,4,5)
vcList<-list(e,f,g)
names(vcList)<-c("e","f","g")
combos<-Reduce(c,lapply(2:length(vcList),
               function(x) combn(1:length(vcList),x,simplify = FALSE))) #possible combos
intr<-lapply(combos, function(x) Reduce(intersect, vcList[x])) #intersect in possible combos
#this!
combos2<-Reduce(c,lapply(2:length(vcList),
               function(x) combn(names(vcList),x, simplify = FALSE))) 
intr2<-lapply(combos2, function(x) Reduce(intersect, vcList[x])) #intersect in possible combos
#names(intr2)<-combos2
names(intr2)<-sapply(combos2, function(i) paste0(i, collapse = "_"))
intr2
str(intr2)
#up to here 
#combos2<-Reduce(combn(names(vcList), 3, simplify = FALSE))

Reduce(c,lapply(2:length(vcList),
               function(x) combn(names(vcList),x, collapse= "_" , simplify = FALSE)))
```

